%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Large Colored Title Article
% LaTeX Template
% Version 1.1 (25/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------

\documentclass[DIV=calc, paper=a4, fontsize=11pt, twocolumn]{scrartcl}	 % A4 paper and 11pt font size

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
\usepackage[english]{babel} % English language/hyphenation
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document

\usepackage{sectsty} % Enables custom section titles
\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands

\usepackage{fancyhdr} % Needed to define custom headers/footers
\pagestyle{fancy} % Enables the custom headers/footers
\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")

\usepackage{enumitem}
\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}

% Headers - all currently empty
\lhead{}
\chead{}
\rhead{}

% Footers
\lfoot{}
\cfoot{}
\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"

\renewcommand{\headrulewidth}{0.0pt} % No header rule
\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule

\usepackage{lettrine} % Package to accentuate the first letter of the text
\newcommand{\initial}[1]{ % Defines the command and style for the first letter
\lettrine[lines=3,lhang=0.3,nindent=0em]{
\color{DarkGoldenrod}
{\textsf{#1}}}{}}

%-----------------------------------------------------------------------------
%	TITLE SECTION
%-----------------------------------------------------------------------------

\usepackage{titling} % Allows custom title configuration

\newcommand{\HorRule}{\color{DarkGoldenrod} \rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title

\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule \fontsize{50}{50} \usefont{OT1}{phv}{b}{n} \color{DarkRed} \selectfont} % Horizontal rule before the title

\title{ABC Information Security Policy} % Your article title

\posttitle{\par\end{flushleft}\vskip 0.5em} % Whitespace under the title

\preauthor{\begin{flushleft}\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{DarkRed}} % Author font configuration

\author{Jordan Ell, } % Your name

\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} % Configuration for the institution name
University of Victoria % Your institution

\par\end{flushleft}\HorRule} % Horizontal rule after the title

\date{} % Add a date here if you would like one to appear underneath the title block

%---------------------------------------------------------------------------

\begin{document}

\maketitle % Print the title

\thispagestyle{fancy} % Enabling the custom headers/footers for the first page 

%-------------------------------------------------------------------------
%	ARTICLE CONTENTS
%--------------------------------------------------------------------------

\section*{Subject Area}

Laptops are personal computers given to ABC employees in order to assist
face paces business process and increase productivity. These laptops
can be used both at work and home to allow employee connectivity to work
at all hours. These laptops can contain all manor of information regarding
ABC, its employees, and customers. Laptops have become a high risk data loss
item because of an increase of malicious software / hardware sophistication
as well as a decrease in these malicious intents cost.

The loss of an ABC issued laptop not only incurs the physical cost of
laptop replacement, but is also subject to data security issues. These data
security issues are also present in unattended laptops as well as stolen. It
is important that ABC personnel become aware of the potential security risks
involved in laptop use and take all possible steps to ensure laptop and data
security of their machine. This policy offers guidance on the user of ABC 
issued laptops in both their physical protection as well as virtual.

\section*{Areas of Concern}

The primary area of concern is the loss of confidential company information
regarding the company itself, employees, or customers. These losses can be
brought on by irresponsibility of laptop ownership which can lead to laptop
theft and potential data theft of the laptop's contents. With newly emerging
quick and relatively cheap forms of data theft such as cold boot 
attacks, laptop
information security has become a primary concern of ABC. Factors of laptop
use which can lead to these attacks and data loss include:
\begin{itemize}
\item Laptop theft.
\item Unattended powered on laptops.
\item Sensitive data being stored on personal laptops.
\item Low refresh rates if DRAM.
\item Unprotected laptop BIOS.
\item Laptops being left on for extended periods of time.
\item Hard drive being unencrypted.
\item Data loss of unsecured laptops and hard drives.
\item DRAM attacks such as cold boot attacks.
\item Company wide security breaches due to a single laptop being attacked.
\end{itemize}

%------------------------------------------------

\section*{Intended Outcome}

This policy for information security is intended to prevent virtual and 
physical
loss to the company in the following ways. First, to prevent physical laptop
theft. Second, to prevent data loss attacks on laptops which need not be stolen
or moved from their location. Finally, to promote responsible laptop use
and ownership among all employees of ABC.

%------------------------------------------------

\section*{Information Security Policy}

The policy outlined here is broken into responsibilities of different types of
personal with a bottom up approach. An employee who falls under any of the listed
categories must strictly adhere to the rules outlined.

%------------------------------------------------

\subsection*{Responsibilities of all Personnel}

\begin{description}
  \item[Technical] \hfill
    \begin{itemize}
    \item Ensure a BIOS password is installed on ABC issued laptop.
    \item Ensure no company sensitive data is stored on issued laptop. This
    includes company, employee and customer related sensitive information.
    \item Ensure hard driver encryption is installed and used on 
    ABC issued laptop.
    \end{itemize}
  \item[Physical] \hfill
    \begin{itemize}
    \item Ensure laptop is locked to immovable object when available.
    \item When laptop is to be left unattended for longer than 10 minutes,
    the laptop must either be locked in a cabinet or other secure location
    or must be powered off while be locked to an immovable object.
    \item When laptop is not being used to an extended period of time (at the
    end of the work day or other), the laptop must be powered off.
    \end{itemize}
  \item[Avoid] \hfill
   \begin{itemize}
    \item Leaving laptop powered on for extended periods of time even in
    a secure location.
    \item Giving laptop to any ABC employee other than IT for an extended
    period of time.
    \end{itemize}
  \item[Report] \hfill
   \begin{itemize}
    \item Any laptop seen unattended and not securely fasten with laptop 
    lock or secured inside locked cabinet. 
    \item Any discrepancy with technical measure such as BIOS password,
    secure data storage, or hard drive encryption.
    \item Any actual and suspected security incidents involving physical
    or data loss.
    \end{itemize}
\end{description}

%------------------------------------------------

\subsection*{Responsibilities of IT}

\begin{description}
  \item[Technical] \hfill
    \begin{itemize}
    \item Ensure only high refresh rate DRAM is purchased and installed
    on any ABC laptop issues to employees.
    \item Ensure any hard drive encryption also flushes DRAM when the
    laptop is powered off.
    \item Ensure proper training in all technical controls are issued to
    new employees of ABC and that compliance metrics are met.
    \end{itemize}
  \item[Compliance] \hfill
    \begin{itemize}
    \item Ensure data security information sessions are held with each
    department of ABC are held at least once a year to update employees
    about new security measures and concerns.
    \item At the department's discretion, perform spot checks of departments
    or employees laptops to ensure policy standards are being met.
    \item Ensure employees properly patch software as needed on laptops as
    patching requires employee presence with BIOS password.
    \end{itemize}
  \item[Avoid] \hfill
   \begin{itemize}
    \item Sensitive data being stored on internet facing
    servers or ensure proper security measure have been taken 
    for data security.
    \end{itemize}
\end{description}
%------------------------------------------------

\subsection*{Responsibilities of HR}

\begin{description}
  \item[Physical] \hfill
    \begin{itemize}
    \item Ensure all new employees are trained on proper laptop security.
    This includes all physical items listed in the responsibilities of all
    personnel.
    \end{itemize}
  \item[Compliance] \hfill
    \begin{itemize}
    \item At the department's discretion, perform spot checks of departments
    or employees laptops to ensure physical policy standards are being met.
    \end{itemize}
\end{description}

\end{document}