%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Trey Hunner (http://www.treyhunner.com/)
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{resume} % Use the custom resume.cls style

\usepackage[left=0.75in,top=0.6in,right=0.75in,bottom=0.6in]{geometry} % Document margins

\name{Jordan Ell} % Your name
\address{2701 Gosworth Rd Apt-303 \\ Victoria, British Columbia, Canada} % Your address
\address{(250)~$\cdot$~415~$\cdot$~9542 \\ jordan.ell7@gmail.com} % Your phone number and email

\begin{document}

I am a Computer Scientist and have been involved in both software development
in industry, as well as computer science and software engineering research
in academia. My broad range of contributions to the feild allow me to tackle
many problems from work flow to architecture to software optimization. I have
also been involved in large scale data security measures and optimizations which
brings experience directly from the feild.

%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Education}

{\bf University of Victoria} \hfill {\em June 2013} \\ 
B.S. in Computer Science with Software Engineering \\
Jamie Cassels Undergraduate Research Scholarship (2012) \\
Presidents Scholarship (2008, 2012) \\
Graduating GPA: 7.73 (3.9)

{\bf University of Victoria} \hfill {\em June 2014} \\ 
MSc. in Computer Science\\
GPA: Unknown at time

\end{rSection}

%----------------------------------------------------------------------------------------
%	WORK EXPERIENCE SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Experience}

\begin{rSubsection}{University of Victoria}{May 2012 - Present}{Software Engineering Researcher}{Victoria, BC}
\item Researching in the realm of socio-technical congruence with applications in awareness among developers.
\item Data mining large scale structured and unstructured data to create technical and social dependency networks
inside of software projects.
\item Creation of bug prediction tools for socially induced failures.
\item Creation of source code analysis tools in order to extract technical dependencies.
\item Applications developed using Java, PostgreSQL, HTML, Javascript, CSS, and AJAX.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{IBM}{September 2010 - April 2011}{Environment Support}{Vancouver, BC}
\item Orchestrated the deployment and maintenance of Linix, Windows, and AIX, project servers.
\item Creation of a single sign on authentication system across Unix and Linux platforms to increase accountability and security.
\item Creation of an OS and middle ware patch management system to track required and applied software patches on servers across multiple levels of security.
\item Creation of automatic tools to ensure server backup systems made adjustments as necessary to complete backup cycles.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{James Evans and Associates}{May 2010 - September 2010}{Oracle Developer}{Victoria, BC}
\item Developed Oracle Forms as a front end for pension management systems.
\item Creation of "Smart Letters" which use an Oracle Database to automatically generate a large set of template letters with stakeholder information.
\item Created and managed large SQL scripts / libraries for maintaining sets of data in a Oracle database.
\end{rSubsection}

\end{rSection}

%----------------------------------------------------------------------------------------
%	COMPETENCIES SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Competencies}


\textbf{Change Leadership} - During my time at IBM, I had led a change in a large portion
of IT infrastructure centred around access management and security. This position
allowed me to create a new vision for our policies and technical tools and follow
through by creating and supporting this vision.\\

\textbf{Analytical Thinking} - During my time at James Evan and Associates, I had
opportunities to organize a database schema. This involved taking a very abstract
view and breaking it down in several smaller and finer grain components to 
create a successful schema.\\

\textbf{Conceptual Thinking} - At IBM, while in charge of data backup procedures, often
backups would fail for non obvious reasons. In order to solve these issues, 
complex event chains would often have to be found in order to address the root
of the problem.\\

\textbf{Long Term Focus} - while implementing a single sign on authentication for 
improved accountability, often staff would complain about a now extra step
needed on their part. The long term focus to avoid these debates was to show
how increased accountability leads to higher security standards and data
integrity.\\

\end{rSection}

%----------------------------------------------------------------------------------------
%	TECHNICAL STRENGTHS SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Knowledge}

\begin{tabular}{ @{} >{\bfseries}l @{\hspace{6ex}} l }
Computer Languages & Java, C++, C\#, C, Python, Ruby \\
Protocols \& APIs & .NET, SDL, OpenGL, XML, JSON, REST \\
Databases & MySQL, PostgreSQL, Oracle Database, DB2 \\
Tools & Git, Svn, CVS, Eclipse, Visual Studio\\
OS & Linux (Ubuntu, Fedora, RedHat), Windows, OSX
\end{tabular}

\end{rSection}

%----------------------------------------------------------------------------------------
%	PROJECTS SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Projects}

All projects availible on GitHub at: github.com/jordanell

\begin{rSubsection}{Impact!}{September 2012 - Present}{Sole Developer}{}
\item An awareness tool used for developers in small to medium size projects written in Java, HTML, CSS, Javascript and AJAX.
\item Provides automatic social recommendations that are based on technical dependencies among developers in the project's source code.
\item Allows developers to subscribe to informational "Streams" which track files, objects, methods and people.
\end{rSubsection}

\begin{rSubsection}{STCharts}{September 2012 - April 2013}{Sole Developer}{}
\item A socio-technical congruence awareness visualization tool written in Java, HTML, CSS, Javascript and AJAX.
\item Developed as part of a fellow researcher's PhD.
\item Allows researchers to input a Git repository as well as a Jira issue tracking URL, which then get analysed for socio-technical congruence over the projects lifetime and is then
displayed using a web front end making use of Google Charts.
\end{rSubsection}

\end{rSection}

%----------------------------------------------------------------------------------------
%	EXAMPLE SECTION
%----------------------------------------------------------------------------------------

%\begin{rSection}{Section Name}

%Section content\ldots

%\end{rSection}

%----------------------------------------------------------------------------------------

\end{document}
